The Concurrent Calculi Formalisation Benchmark

### 2024-04-17 short-paper/short-paper.tex
minor cosmetic improvements

Our benchmark challenges (detailed on our website linked below) consider
\emph{in isolation} three key aspects that frequently pose difficulties
when mechanising concurrency theory, which we will discuss in more detail in the next section: \emph{linearity}, \emph{scope
  extrusion}, and \emph{coinductive reasoning}.
Mechanisations must often address several of these aspects at the same time; however, we
see the combination of techniques as a next
step, as discussed in \cref{sec:going-beyond}.

@@

\subsection{Linearity}
Linear typing systems allow for the tracking of resource usage in a program. In
the case of typed (in particular, session-typed) process calculi, linearity is
widely used for checking if and how a channel is used to send or receive values.
This substructurality~\cite[Ch. 1]{Pierce2004-oq} gives rise to mechanisation difficulties: \eg deciding how to \emph{split the
typing context} in a parallel composition.

@@

Higher-order abstract syntax (HOAS) has seen extensive
use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010,ChaudhuriCM15,Castro-Perez2021}. Its
weak form aligns well with mainstream inductive proof assistants,
significantly simplifying the encoding of typing systems and operational
semantics. However, when addressing more intricate concepts like
bisimulation, extensions to HOAS are needed. These extensions may take
the form of additional axioms~\cite{Honsell2001} or require niche proof assistants
such as Abella, which features a special quantifier for handling
properties related to names~\cite{GacekMN11}.

@@

Our coinduction challenge only addresses two notions of process equivalence, but many more exist in the literature.
Coinduction may also play a role in recursive processes and session types: recursive session types can be expressed in \emph{infinitary form} by interpreting their typing rules coinductively~\cite{DerakhshanPfenning22,HornePadovani23}.

>>>>

Our benchmark challenges (detailed on our website linked below)
consider \emph{in isolation} three key aspects that frequently pose
difficulties when mechanising concurrency theory: \emph{linearity},
\emph{scope extrusion}, and \emph{coinductive reasoning}, as we will
discuss in more detail in the next section.  Mechanisations must often
address several of these aspects at the same time; however, we see the
combination of techniques as a next step, as argued in
\cref{sec:going-beyond}.

@@

\subsection{Linearity}
Linear typing systems enables the tracking of resource usage in a program. In
the case of typed (in particular, session-typed) process calculi, linearity is
widely used for checking if and how a channel is used to send or receive values.
This substructurality~\cite[Ch. 1]{Pierce2004-oq} gives rise to mechanisation difficulties: \eg deciding how to \emph{split the
typing context} in a parallel composition.

@@

Higher-order abstract syntax (HOAS) has seen extensive
use in formal reasoning in this area~\cite{Despeyroux2000,Honsell2001,Tiu2010,ChaudhuriCM15,Castro-Perez2021}. Its
weak form aligns reasonably well with mainstream inductive proof assistants,
significantly simplifying the encoding of typing systems and operational
semantics. However, when addressing more intricate concepts like
bisimulation, extensions to HOAS are needed. These extensions may take
the form of additional axioms~\cite{Honsell2001} or require niche proof assistants
such as Abella, which features a special quantifier for handling
properties related to names~\cite{GacekMN11}.

@@

Our coinduction challenge only treats two notions of process equivalence, but many more exist in the literature.
Coinduction may also play a role in recursive processes and session types: recursive session types can be expressed in \emph{infinitary form} by interpreting their typing rules coinductively~\cite{DerakhshanPfenning22,HornePadovani23}.
